```{r {{id}}-setup, echo=FALSE}
#A species is randomly sampled if this template is run interactively.
#Else, the id is equal to the {{id}} that is passed through knit_expand
if (interactive()) {
  this_par <- sample(unique(colnames(data_cv)[6:39]), 1)
} else {
  this_par <- "{{id}}"
}
data_cv_all_2 <- data_cv %>%
  mutate(binned = cut_number(x = data_cv %>% dplyr::pull(this_par), n = 4)) %>%
  dplyr::select(`ID set`, binned) %>%
  right_join(data_cv_all_1, by = join_by(`ID set` == ID))
# selection <- iris %>%
#   dplyr::filter(Species == this_id)
```

#### Analysis of the `r stringr::str_to_title(this_par)` parameter

The `r this_par` parameter takes values between `r min(data_cv %>% dplyr::pull(this_par))` and `r max(data_cv %>% dplyr::pull(this_par))`. `r this_par` is defined as "`r parameters %>% dplyr::filter(Parameter == this_par) %>% dplyr::pull(Info)`". It is split up in four equal-sized discrete bins (same number of instances in each bin). We now compare the run times for instances in each of the four bins for both @watermeyer2022partition and @creemers2021.

Figure \@ref(fig:{{id}}-KM) shows the different Kaplan- Meier curves.

```{r {{id}}-KM, fig.cap="Comparing the run time for WZ (left) and C (right). Note that the x-axis does not match", echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 5}
p1 <- survfit2(Surv(Time, status) ~ binned,
         data = data_cv_all_2 %>% filter(author == "WZ")) %>%
  ggsurvfit() +
  labs(
    x = "milliseconds",
    y = "Proportion of instances not solved to optimality"
    ) + 
  ylim(0, 1) +
  #xlim(0, max(data_cv_all_1$Time))
  scale_color_manual(name = this_par,
                     values = colorRampPalette(colors = c("#99CCFF", "#003366"))(4))
p2 <- survfit2(Surv(Time, status) ~ binned,
         data = data_cv_all_2 %>% filter(author == "C")) %>%
  ggsurvfit() +
  labs(
    x = "milliseconds",
    y = "Proportion of instances not solved to optimality"
    ) + 
  ylim(0, 1) +
  #xlim(0, max(data_cv_all_1$Time))
  scale_color_manual(name = this_par,
                     values = colorRampPalette(colors = c("#99CCFF", "#003366"))(4))

combined <- p1 + p2 & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect")
```

