```{r {{id}}-setup, echo=FALSE}
#A species is randomly sampled if this template is run interactively.
#Else, the id is equal to the {{id}} that is passed through knit_expand
if (interactive()) {
  this_par <- sample(unique(colnames(data_cv)[6:39]), 1)
} else {
  this_par <- "{{id}}"
}

data_cv_all_2 <- data_cv %>%
  mutate(binned = chop_equally(x = data_cv %>% dplyr::pull(this_par),
                               groups = 4),
         continuous = data_cv %>% dplyr::pull(this_par)) %>%
  dplyr::select(IDset, binned, continuous) %>%
  right_join(data_cv_all_1 %>% filter(author == "C"),
             by = join_by(IDset == ID))

data_psplib_all_2 <- data_psplib[data_psplib[,this_par] != "n.f.", ] %>%
  mutate(
    binned = chop_equally(
      x = data_psplib[data_psplib[,this_par] != "n.f.", ] %>%
        dplyr::pull(this_par), groups = 4),
    continuous = data_psplib[data_psplib[,this_par] != "n.f.", ] %>%
      dplyr::pull(this_par)) %>%
  dplyr::select(IDset, binned, continuous) %>%
  right_join(data_psplib_all_1 %>% filter(author == "C"),
             by = join_by(IDset == ID)) %>%
  filter(!is.na(binned)) 
nb_groups <- length(unique(data_cv_all_2$binned)) #should be 4 but chop_equally adjusts if necessary
# selection <- iris %>%
#   dplyr::filter(Species == this_id)
```

#### Analysis of the `r stringr::str_to_title(this_par)` parameter

The `r this_par` parameter takes values between `r min(data_cv %>% dplyr::pull(this_par))` and `r max(data_cv %>% dplyr::pull(this_par))`. `r this_par` is defined as "`r parameters %>% dplyr::filter(Parameter == this_par) %>% dplyr::pull(Info)`". It is split up in `r nb_groups` equal-sized discrete bins (same number of instances in each bin). We now compare the run times of @creemers2021 for instances in each of the `r nb_groups` bins for both the PSPLIB and CV dataset.
Figure \@ref(fig:{{id}}-KM) shows the different Kaplan-Meier curves.

```{r {{id}}-KM, fig.cap="Comparing the run time for Creemers (2021) for the J60 (left) and CV (right) instance sets.", echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 4}
p1 <- survfit2(Surv(Time, status) ~ binned,
         data = data_psplib_all_2 %>% filter(author == "C")) %>%
  ggsurvfit() +
  labs(
    x = "milliseconds",
    y = "Proportion not solved optimally"
    ) + 
  ylim(0, 1) +
  #xlim(0, max(data_cv_all_1$Time))
  scale_color_manual(name = this_par,
                     values = colorRampPalette(colors = c("#99CCFF", "#003366"))(4))
p2 <- survfit2(Surv(Time, status) ~ binned,
         data = data_cv_all_2 %>% filter(author == "C")) %>%
  ggsurvfit() +
  labs(
    x = "milliseconds",
    y = "Proportion not solved optimally"
    ) + 
  ylim(0, 1) +
  #xlim(0, max(data_cv_all_1$Time))
  scale_color_manual(name = this_par,
                     values = colorRampPalette(colors = c("#99CCFF", "#003366"))(4))

combined <- p1 + p2 & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect")

data_both <- rbind(data_cv_all_2 %>% filter(author == "C"),
                   data_psplib_all_2 %>% filter(author == "C"))

library("survival")
if(this_par != "LE") {
res.cox.bin <- coxph(Surv(Time, status) ~ binned,
                 data =  data_both)
test.ph.bin <- cox.zph(res.cox.bin)
res.cox.cntn <- coxph(Surv(Time, status) ~ continuous,
                 data = data_both)
test.ph.cntn <- cox.zph(res.cox.cntn)
}
```

```{r conditional_print, child='_text_PH.Rmd', eval = (this_par != "LE"), result = 'asis'}
```

```{r conditional_print, child='_text_no_PH.Rmd', eval = (this_par == "LE"), result = 'asis'}
```

```{r {{id}}-PHassumption, fig.cap="Schoenfeld residuals for the binned and continuous version of the complexity parameter.", echo = FALSE, message = FALSE, warning = FALSE, fig.width = 9, fig.height = 4, eval = (this_par != "LE")}
par(mfrow = c(1,2))
plot(test.ph.bin)
plot(test.ph.cntn)
```

\clearpage